Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario

### 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Added data files from TTS with cross-tab of age by income and edited the data processing to read them

I downloaded a cross-tabulation of travel mode by household income by traffic analysis zone. Read the shape file with the traffic analysis zone boundaries:
```{r}
tts_taz <- st_read("input-data-files/tts2016_2006zn_2018_region.shp")
```

@@

@@

@@

Filter age:
```{r}
travel_modes_1 <- travel_modes %>%
  dplyr::filter(income >= 1 & income <= 2)
```

@@

@@

@@

Save data files:
```{r}
save(data_da_2016, file = "output-data-files/data_da_2016.RData", compress = "xz")
save(dwelling_network_points_2016, file = "output-data-files/dwelling_network_points_2016.RData", compress = "xz")
save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
save(modes_less40k, file = "output-data-files/modes_less40k.RData", compress = "xz")
save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")
```

>>>>

We downloaded a cross-tabulation of travel mode by household income by traffic analysis zone. We also downloaded a cross-tabulation of age of person by household income by traffic analysis zone.

Read the shape file with the traffic analysis zone boundaries:
```{r}
tts_taz <- st_read("input-data-files/tts2016_2006zn_2018_region.shp")
```

@@

Read population by age by income:
```{r}
age_income <- read_csv("input-data-files/ttsCross65713.csv",
                         col_types = cols(gta06_hhld = col_double(),
                                          age = col_double(),
                                          population = col_double(),
                                          income = col_double()
                                          ))
```

@@

Filter low income:
```{r}
age_income <- age_income %>%
  dplyr::filter(income <= 2)
```

@@

Filter income:
```{r}
travel_modes_1 <- travel_modes %>%
  dplyr::filter(income >= 1 & income <= 2)
```

@@

Summarize population by age by income by traffic analysis zone. Children:
```{r}
low_income_children <- age_income %>%
  dplyr::filter(age <= 18) %>%
  rename(TAZUID = gta06_hhld) %>%
  group_by(TAZUID) %>%
  summarize(children = sum(population),
            .groups = "drop")
```

Adults:
```{r}
low_income_adults <- age_income %>%
  dplyr::filter(age > 18 & age < 65) %>%
  rename(TAZUID = gta06_hhld) %>%
  group_by(TAZUID) %>%
  summarize(adults = sum(population),
            .groups = "drop")
```

Seniors:
```{r}
low_income_seniors <- age_income %>%
  dplyr::filter(age >= 65) %>%
  rename(TAZUID = gta06_hhld) %>%
  group_by(TAZUID) %>%
  summarize(seniors = sum(population),
            .groups = "drop")
```

Join tables with population by age and add geometry:
```{r}
population_age_income <- tts_taz %>%
  transmute(TAZUID = gta06, 
            geometry) %>%
  left_join(low_income_children,
            by = "TAZUID") %>%
  left_join(low_income_adults,
            by = "TAZUID") %>%
  left_join(low_income_seniors,
            by = "TAZUID") %>%
  mutate(children = replace_na(children, 0),
         adults = replace_na(adults, 0),
         seniors = replace_na(seniors, 0))
```

@@

## Before saving

@@

Save data files:
```{r}
save(data_da_2016, file = "output-data-files/data_da_2016.RData", compress = "xz")
save(dwelling_network_points_2016, file = "output-data-files/dwelling_network_points_2016.RData", compress = "xz")
save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
save(modes_less40k, file = "output-data-files/modes_less40k.RData", compress = "xz")
save(population_age_income, file = "output-data-files/population_age_income.RData", compress = "xz")
save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")
```

### 2021-04-20 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Added data files from TTS with cross-tab of age by income and edited the data processing to read them

It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial]. However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks. Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource (e.g. Enns et al. [-@enns2020experiences]), we are not aware of any research that has focused on the capturing the geographic component of accessibility.

The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring]. Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery in Vancouver [@holmes2018nothing]. However, as with research into the prevalence of food deserts, accessibiltiy to food banks is likely not evenly distributed throughout a city with variations in access attributable to transportation network characteristics and the spatial distribution of food bank locations. Furthermore, policy responses to the COVID-19 pandemic add to the distress of vulnerable households. Non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which they are more likely to be reliant [e.g., @deweese2020tale]. At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns around the delivery of service in safe conditions and possible cancellation of food service programs.

@@

Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake (Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009). This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage (Bazerghi, McKay, and Dunn 2016). Previous studies have aimed to look at the relationship between the built food environment and sociodemographic characteristics with qualitative and quantitative, or a mixed -method approach.

@@

Methods
===================

>>>>

It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial]. However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks. Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.

The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring]. Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery in Vancouver [@holmes2018nothing]. However, as with research into the prevalence of food deserts, accessibility to food banks is likely not evenly distributed throughout a city with variations in access attributable to transportation network characteristics and the spatial distribution of food bank locations - as well as the population that likely needs them. Furthermore, policy responses to the COVID-19 pandemic add to the distress of vulnerable households. Non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which low income populations are more likely to be reliant [e.g., @deweese2020tale]. At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.

@@

Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake (Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009). This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage (Bazerghi, McKay, and Dunn 2016). Previous studies have aimed to look at the relationship between the built food environment and socio-demographic characteristics with qualitative and quantitative, or a mixed -method approach.

@@

Data and Methods
===================

## Data
